---
title: "Austrian presidential elections manipulated? &nbsp; Definitely not: Charts and Probabilities"
author: "Erich Neuwirth"
date: "August 27, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

## Load packages

```{r packages}
library(stringr)
library(ggplot2)
library(dplyr)
library(intubate)
library(magrittr)
```
## Get data and compute values needed

```{r load_data}
load(file=file.path("Rdata","workdata.RData"))
```

```{r compute_percentages}
workdata %<>%
  mutate(Hofer_ant_u=Hofer_u/Tot_u,
         Hofer_ant_m=Hofer_m/Tot_m) 

```


```{r total_results}
workdata %>% select(Hofer_u,Hofer_m,Tot_u,Tot_m) %>%
  colSums() -> toterg
shiftneeded <- ceiling((toterg[3]+toterg[4])/2-(toterg[1]+toterg[2]))

```

```{r separate_results}
workdata %>%
  select(incrim11,Tot_m,Hofer_m,Tot_u,Hofer_u) %>%
   group_by(incrim11) %>%
  summarise_each("sum") ->
  sums11
tot11m <- sums11[[1,2]]
hofer11m <- sums11[[1,3]]
hofer11u <- sums11[[1,5]]
hoferneeds11 <- (hofer11m + shiftneeded)
```


## Charts

### Without modifications

```{r real_chart}
mymax <- with(workdata,max(Hofer_ant_m,Hofer_ant_u)+0.075)
workdata %>%
  mutate(incrim=ifelse(incrim11=="beanst.","contaminated","clean")) %>%
  ggplot(aes(x=Hofer_ant_u,y=Hofer_ant_m,size=Tot_m,color=incrim)) +
  geom_point() +
  geom_abline(slope=1,intercept=0,linetype="dotted") +
  scale_size_continuous("Wahlkarten-\nstimmen",range=c(1,3),guide=FALSE) +
  scale_color_manual("court\nclassification",values=c("limegreen","red")) +
  scale_x_continuous("Hofer ballot votes",labels=scales::percent,
                     limits=c(0,mymax)) +
  scale_y_continuous("Hofer mail votes",
                     labels=scales::percent,limits=c(0,mymax)) ->
  pic1
pic1
```

```{r}
png("pic1.png",width=700,height=600)
pic1
dev.off()
system("open pic1.png")
```

### German labels, Villach marked

```{r real_chart_german}
villachdf <- workdata %>% 
  filter(GKZ=="G20200")

mymax <- with(workdata,max(Hofer_ant_m,Hofer_ant_u)+0.075)
workdata %>%
  mutate(incrim=ifelse(incrim11=="beanst.","unsauber","sauber")) %>%
  ggplot(aes(x=Hofer_ant_u,y=Hofer_ant_m,size=Tot_m,color=incrim)) +
  geom_point() +
  geom_abline(slope=1,intercept=0,linetype="dotted") +
  geom_point(aes(x=Hofer_ant_u,y=Hofer_ant_m),data=villachdf,size=5,shape="O",inherit.aes=FALSE) +

  scale_size_continuous("Wahlkarten-\nstimmen",range=c(1,3),guide=FALSE) +
  scale_color_manual("VfGH\nKlassifikation",values=c("limegreen","red")) +
  scale_x_continuous("Hofer Urnenwahl",labels=scales::percent,
                     limits=c(0,mymax)) +
  scale_y_continuous("Hofer Wahlkarten",
                     labels=scales::percent,limits=c(0,mymax)) ->
  pic1d
pic1d
```


```{r}
pdf("pic1d.pdf",width=6,height=4.5)
pic1d
dev.off()
system("open pic1d.pdf")
```

### Modifications in 11 districts


```{r shifted_chart}
mymax <- with(workdata,max(Hofer_ant_m,Hofer_ant_u)+0.075)

workdata %>%
  mutate(incrim=ifelse(incrim11=="beanst.","contaminated","clean")) %>%
  mutate(incrim=factor(incrim,levels=c("clean","dubious","contaminated"))) %>%
  mutate(manipval=ifelse(incrim11=="beanst.",shiftneeded*Tot_m/tot11m,0)) %>%
  mutate(hh=(Hofer_m+manipval)/Tot_m) %>%
  ggplot(aes(x=Hofer_ant_u,y=hh,
             size=Tot_m,color=incrim)) +
  geom_point() +
  geom_abline(slope=1,intercept=0,linetype="dotted") +
  scale_size_continuous("Wahlkarten-\nstimmen",range=c(1,4),guide=FALSE) +
  scale_color_manual("court\nclassification",values=c("limegreen","red")) +
  scale_x_continuous("Hofer ballot votes",labels=scales::percent,
                     limits=c(0,mymax)) +
  scale_y_continuous("Hofer mail votes",
                     labels=scales::percent,limits=c(0,mymax)) ->
  pic11
pic11
```



```{r}
png("pic11.png",width=700,height=600)
pic11
dev.off()
system("open pic11.png")
```

### Modifications in 11 districts, German


```{r}
mymax <- with(workdata,max(Hofer_ant_m,Hofer_ant_u)+0.075)

workdata %>%
  mutate(incrim=ifelse(incrim11=="beanst.","unsauber","sauber")) %>%
  mutate(manipval=ifelse(incrim11=="beanst.",shiftneeded*Tot_m/tot11m,0)) %>%
  mutate(hh=(Hofer_m+manipval)/Tot_m) %>%
  ggplot(aes(x=Hofer_ant_u,y=hh,
             size=Tot_m,color=incrim)) +
  geom_point() +
  geom_abline(slope=1,intercept=0,linetype="dotted") +
  
  
  scale_size_continuous("Wahlkarten-\nstimmen",range=c(1,3),guide=FALSE) +
  scale_color_manual("VfGH\nKlassifikation",values=c("limegreen","red")) +
  scale_x_continuous("Hofer Urnenwahl",labels=scales::percent,
                     limits=c(0,mymax)) +
  scale_y_continuous("Hofer Wahlkarten",
                     labels=scales::percent,limits=c(0,mymax)) ->
  pic11d
pic11d
```


```{r}
pdf("pic11d.pdf",width=6,height=4.5)
pic11d
dev.off()
system("open pic11d.pdf")
```




## Computing probabilities for changing election outcome in different scenarios

### Linear model with 11 districts removed

```{r linear_model}
workdata %>% filter(!(incrim11 == "beanst.")) %>%
  ntbt_lm(Hofer_m ~ Hofer_u - 1,weights=1/Tot_m) ->
  model11
```

### Data frame for prediction sum of 11 districts

```{r predictdf}
preddf11 <- data.frame(Hofer_u=hofer11u,Tot_m=tot11m)  
```

### Auxiliary function

`predict.lm` computes prediction intervals for given confidence levels.
Our problem is the other way round, we compute the probability od
the prediction exceeding a given target value.

To do this, we compute a prediction interval for 95% level using
the classical prediction estimator for linear models build into R
(which uses a rescaled t-disribution),
extract the value by which the t-distribution for this prediction has been rescaled,
and use this scale value to compute the probability for a given prediction interval width. 


```{r prob_function}
find_prob_pred_geq <- function(model,preddf,targetval){
  pred <- predict.lm(model,preddf,interval="prediction",weights=~1/Tot_m,level=0.95)
  predscale <- (pred[1]-pred[2])/qt(0.975,df.residual(model))
  res <- 1-pt((targetval-pred[1])/predscale,df.residual(model))
  names(res) <- NULL
  res
}
```


## Probability for sum of 11 districts changing election outcome, model without 11 districts

```{r prob_value}
myprob <- find_prob_pred_geq(model11,preddf11,hoferneeds11)  
myprob %>% cat()
```

```{r print_helper}
printfixed <- function(x)sprintf("%.0f",x)
```

We can check that the the probability really gives the target value as limit
of the confidence interval.

Hofer has `r printfixed(hofer11m)` votes in the contaminated districts.
He needs `r printfixed(shiftneeded)` more mail votes there, making for a total of
`r printfixed(hoferneeds11)` mail votes in the contaminated districts.

Let us check the corresponding prediction confidence interval with probability `r sprintf("%.12f",1-2*myprob)`

```{r final check}
predict(model11,preddf11,interval="prediction",weights=~1/Tot_m, level=1-2*myprob)
```

